{\LARGE \textsc{Structural controlability of pollination networks}}

### 2016-05-02 paper/manuscript.Rnw
update methods

To find the maximum matching in a pollination network we used an alternative bipartite representation of the network (\autoref{fig:explain_mm}b) in which the levels indicate the outgoing and incoming links to each node. Finding a matching in this alternative representation is equivalent to finding the set of links such that each node on the top level is matched to at most one node on the bottom level and vice versa \autocite{West2001, Csardi2006a}.

The link between network controllability and the maximum matching also has an intuitive basis. In control theory, a network is fully controlable when each node in the network has a "superior", or input that drives its dynamics. Matched nodes can be controled within the network by the node pointing to it---its superior. However, unmatched nodes have no superior in the network and must be directly controlled by external inputs \autocite{Liu2011}, what we refer to as management interventions .

Systems that require a large number of inputs are more difficult to control. The \textit{manageability} of a community is therefore determined by the minimum number of species on which we would need to apply management interventions if we were to gain full control of the abbundances of all species in the ecological community ($n_d$, \autoref{eq:m}).

\begin{equation} \label{eq:m}
m = 1- \frac{n_d}{s}
\end{equation}

We quantified the manageability of ten paired plant-pollinator communities. Each pair was composed by a community invaded by a plant and a community effectively free of the invasive species (Table S1). The networks were constructed from previously published visitation data collected from pollination communities in Bristol, UK \autocite{Lopezaraiza-Mikel2007} and Cap de Creus National Park, Spain \autocite{Bartomeus2008}. In the four British uninvaded communities the non-invaded plots were obtained by experimentally removing all the flowers of invasive species \textit{Impatients grandulifera}. In contrast, the Spanish uninvaded communities were obtained from plots that had not yet been colonised by the invasive species \textit{Carpobrotus affine acinaciformis} or \textit{Opuntia stricta}.

The analysed networks were quantified using visitation frequency, which has been shown to be an appropriate surrogate for inter-specific effects in pollination networks \autocite{Vazquez2005, Bascompte2006}. Nevertheless we evaluated the effect of using pollinator efficiency or importance as measures of species interactions \autocite{NeEman2010, Ballantyne2015} and found consistent results (see Supplementary Information).

\subsubsection*{Interspecific effects}

Our maximum matching algotithm requires a directed network in which a link between species $i$ and $j$ pointing to species $j$ indicates that the abundance of species $j$ can be affected by the abundance of species $i$. The extent to which a pair of interacting species affect each other---the mutualism strengtht---can be quantified by the magnitude of the mutual dependence \autocite{Bascompte2006}. Therefore, we initially assumed that a pair of interacting species can affect each other abundances. The effect on each other being inverselly proportional to the magnitude of the dependencies. This assumption renders a bipartite network in which interacting pairs are connected by two directed links (\autoref{fig:explain_cd}b).

@@

Nevertheless, mutual dependencies have been shown to be highly assymetric in natural communities \autocite{Bascompte2006}. In other words, if a plant species is largely dependent on a pollinator species, then the pollinator tends to depend only very weakly on the plant. This suggest that it may be possible to relax our initial assumption and simplify the network such that interacting species are connected by one directed link when mutual dependencies are assymetric. The direction of this link being consistent with the dominant direction of control in the bidirectional case (\autoref{fig:explain_cd}c). The extent to which a species $i$ affects species $j$ relative to the extent to which species $j$ affects species $i$ can then be summarised by the interaction asymmetry (\autoref{eq:a}).

\begin{equation} \label{eq:a}
a(i,j) = \frac{\left | d_{ij}-d_{ji} \right |}{\textup{max}\left ( d_{ij}, d_{ji} \right )}
\end{equation}

We quantified the manageability of the analysed pollination communities by finding the maximum matching of the corresponding pollination network following the method previously explained. If a network had more than one component (two species are in different components if there exist no path between them) we analysed the largest one. We then performed a paired Spearman correlation test to compare the manageability obtained when including bidirectional effects between interacting species, and when simplyfing the network to include only the dominant direction of the asymmetry.

@@

The manageability of a community $m$ provides an indication of the magnitude of $n_d$ but not the identity of the species that can compose this set. In fact, maximum matchings are often not unique in ecological networks and different sets of species of size $n_d$ could be used to control the network \autocite{Liu2011}. Nevertheless, species vary in terms of the frequency with which they are classified as superior nodes (\autoref{fig:explain_mm}). This frequency provides an indication of their relative importance to drive the state of the community. We therefore call \textit{driver species} those that are identified as needing directed interventions in all possible matchings\todo{elaborate concept}, and consequently likely to have a disproportionate effect on the abbundances of other species in the community.

Unlike unweighted networks, where a maximum matching is defined by its size (the number of matched edges), in weighted networks, a maximum matching is one whose edges have the largest possible total weight among all possible matchings. This means that in weighted networks there is a number of matchings that have maximal size, but are not maximum matchings because its total weight is not maximal too. However the weight of species dependencies is unexact, a reflection of the fact that interaction strengths and sampling methods themselves are stochastic to some extent.

We therefore identified the driver species in each of the networks by computing all possible matchings in the network that have the same size and have a weight above a cut-off of 0.5 times the weight of the maximum matching\todo{justify}. To do so, we first generated the network's line graph---an alternative representation in which nodes are the interactions of the original network and links represent common ending points. We then calculated the complement of the line graph---a graph in which nodes are adjecent if and only if they are not adjancent in the line graph. All possible maximum matchings can be found by enumerating all cliques of size $1-n_d$ in the complement graph \autocite{West2001, Csardi2006a}. Because of the number of matchings depends on the cut-off and the weighting of the links, we evaluated the effects of i) using a different cut-offs value between 0 and 1 and ii) weighting the links by the number of visits.

>>>>

Given a maximum matching, the link to network controllability has a relatively straightforward and intuitive basis. In control theory, a network is fully controlable when each node has a "superior", this is, an input that drives its dynamics. Matched nodes can thus be controled by the node pointing to it---its superior. Unmatched nodes on the other hand, have no superior in the network and must be directly controlled by external inputs \autocite{Liu2011}. In an ecological context, this is what we refer to as management interventions.

Systems that require a large number of external inputs are more difficult or costly to control. The \textit{manageability} of a community $m$ is therefore determined by the proportion $n_d$ of species on which we would need to apply management interventions in order to gain full control of the abundances of all species in the ecological community, $$m = 1- \frac{n_d}{s},$$ where $s$ is the total number of species in the community.

\subsubsection*{Weighting links}

Recall that our maximum matching algorithm requires a directed network in which a link between species $i$ and $j$ pointing to species $j$ indicates that the abundance of $j$ can be affected by the abundance of $i$. In ecological networks, establishing a directionality for the links between species is relatively straightforward when they represent resources or energy flow, however, interspecific effects---particularly in pollination networks---are not strictly directed. Nevertheless, the extent to which a pair of interacting species affect each other can be quantified by the magnitude of the mutual dependence \autocite{Bascompte2006}. As the dependencies are bidirectional, this assumption renders a bipartite network in which interacting pairs are connected by two directed links (\autoref{fig:explain_cd}b).

@@

Nevertheless, mutual dependencies have been shown to be highly assymetric in natural communities \autocite{Bascompte2006}. In other words, if a plant species is largely dependent on a pollinator species, then the pollinator tends to depend only very weakly on the plant. This suggest that it may be possible to relax our initial assumption and simplify the network such that interacting species are connected by one directed link when mutual dependencies are highly assymetric (\autoref{fig:explain_cd}c). The extent to which a species $i$ affects species $j$ relative to the extent to which species $j$ affects species $i$ can then be summarised by the interaction asymmetry.

$$a(i,j) = \frac{\left | d_{ij}-d_{ji} \right |}{\textup{max}\left ( d_{ij}, d_{ji} \right )}$$

\subsubsection*{Relative species importance}

\todo[inline]{insert ecological motivation}

The manageability of a community $m$ provides an indication of the magnitude of $n_d$ but not the identity of the species that compose this set. Moreover, maximum matchings are often not unique in for a given network and different sets of species of size $n_d$ could be used to control the network \autocite{Liu2011}. Species vary in terms of the frequency with which they are classified as superior nodes (\autoref{fig:explain_mm}), and this frequency provides an indication of their relative importance in driving the state of the community.

We therefore call \textit{driver species} those that are identified as being a superior node in all\todo{check} possible matchings\todo{elaborate concept}, and consequently likely to have a disproportionate effect on the abbundances of other species in the community. Unlike unweighted networks, where a maximum matching is defined by its matching size (the number of matching links), a maximum matching in a weighted network is one whose links have the largest possible total weight among all possible matchings. This means that in weighted networks there is a number of matchings that have maximal size, but are not maximum matchings because their total weight is not maximal.

We therefore identified the driver species in each of the networks by computing all possible matchings in the network that have the same maximal size while allowing for some tolerance regarding the weight. We relax the matching weight requirement because empirical intreaction strengths, from which we calculate dependencies, are stochastic to some extent and depend on the sampling method and intensity. To compute all the maximal matchings, we first generated the network's line graph---an alternative representation in which nodes are the interactions of the original network and links represent common ending points. We then calculated the complement of the line graph---a graph in which nodes are adjacent if and only if they are not adjacent in the line graph. All possible maximum matchings can be found by enumerating all cliques of size $1-n_d$ in this complement graph \autocite{West2001, Csardi2006a}.

\subsection*{Empirical Data}

In the last section we outlined the theoretical and computational underpinings of our approach. Now we illustrate its empirical application using ten pairs of plant-pollinator communities for which we calculated their manageability and the relative importance of species. Each network pair was composed by a community invaded by a plant and a community effectively free of the invasive species (Table S1). The networks were constructed from previously published visitation data collected from pollination communities in Bristol, UK \autocite{Lopezaraiza-Mikel2007} and in Cap de Creus National Park, Spain \autocite{Bartomeus2008}. In the four British uninvaded communities, the non-invaded plots were obtained by experimentally removing all the flowers of the invasive species \textit{Impatients grandulifera}. In contrast, the Spanish uninvaded communities were obtained from plots that had not yet been colonised by either of the invasive species \textit{Carpobrotus affine acinaciformis} or \textit{Opuntia stricta}.

The structure of these networks was determined using visitation frequency which has been shown to be an appropriate surrogate for inter-specific effects in pollination networks \autocite{Vazquez2005, Bascompte2006}. To examine the importance of this decision, we also evaluated the effect of using pollinator efficiency or importance as alternative measures of species interactions \autocite{NeEman2010, Ballantyne2015}, and we found consistent results for all of these options(see Supplementary Information).

We started by quantifying the manageability of the analysed pollination communities. To do that, we calculated the maximum matching of the corresponding pollination network, and estimated the minimum number of species that need directed intervecntions to fully control the species abbundances in the community. If a network had more than one component (two species are in different components if there exist no path between them) we analysed the largest one. In order to check that

We then performed a paired Spearman correlation test to compare the manageability obtained when including bidirectional effects between interacting species, and when simplyfing the network to include only the dominant direction of the asymmetry.

@@

 However the weight of species dependencies is unexact, a reflection of the fact that interaction strengths and sampling methods themselves are stochastic to some extent.

We therefore identified the driver species in each of the networks by computing all possible matchings in the network that have the same size and have a weight above a cut-off of 0.5 times the weight of the maximum matching\todo{justify}.

Because of the number of matchings depends on the cut-off and the weighting of the links, we evaluated the effects of i) using a different cut-offs value between 0 and 1 and ii) weighting the links by the number of visits.
